### 2022-04-25 old-papers/draft2/main.tex
Response to Julien's comments on draft 2

\documentclass[fleqn,usenatbib]{mnras}

\usepackage[T1]{fontenc}

\DeclareRobustCommand{\VAN}[3]{#2}
\let\VANthebibliography\thebibliography
\def\thebibliography{\DeclareRobustCommand{\VAN}[3]{##3}\VANthebibliography}

%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{colortbl}
\usepackage{dsfont}
\usepackage{siunitx}
\usepackage{bm}

% For the sideways figure
\usepackage{pdflscape}
\usepackage{afterpage}

\usetikzlibrary{arrows.meta}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
\newcommand{\jtd}[1]{ {\bf{\color{red} JTD: #1}} }
\newcommand{\unit}[1]{\bm{\hat{#1}}}

\newcommand{\parens}[1]{\left( #1 \right)}
\newcommand{\brackets}[1]{\left[ #1 \right]}
\newcommand{\quat}[1]{\widetilde{\bm{#1}}}

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

\title[Flyby Constraints on Asteroids Interiors]{Constraining the Interiors of Asteroids Through Close Encounters}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[Jack T. Dinsmore, Julien de Wit]{
Jack T. Dinsmore,$^{1}$\thanks{E-mail: jtdinsmo@mit.edu}
Julien de Wit$^{2}$
\\
% List of institutions
$^{1}$Department of Physics, Massachusetts Institute of Technology\\
$^{2}$Department of Earth, Atmospheric, and Planetary Science, Massachusetts Institute of Technology
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2022}

% Don't change these lines
\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

% Abstract of the paper
\begin{abstract}
  Knowledge of the interior distribution of an asteroid can reveal its composition and constrain its evolutionary history. However, most asteroid observational techniques are not sensitive to interior structure. We investigate the degree to which interior density distributions can be extracted from angular velocity data gathered during a close encounter. We derive the equations of motion for a rigid asteroid's orientation and angular velocity to arbitrary order and use them to generate synthetic angular velocity data for a representative asteroid on a close Earth encounter. Using Markov Chain Monte Carlo fits, we re-extract the density moments of the asteroid in a wide range of scenarios to measure the degree to which best-fitting parameter precision is affected. Specifically, we use many injection-retrieval tests to study fit precision's dependence on the asteroid's moment of inertia, observational precision and cadence, orbital parameters, and initial spin pole direction as well as the planet being encountered, the quantity of near-pericentre data, and the central body oblateness. Finally, we discuss the degeneracy between the density moments and the actual density distribution and propose three models to construct a representative density distribution from fit results.
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
  minor planets, asteroids: general -- methods: data analysis
\end{keywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%

\section{Introduction}

Over the past twenty years, the increase in quantity and quality of sensitive all-sky surveys has prompted the discovery of numerous asteroids. Such advances have been made via ground-based surveys such as the Catalina Sky Survey \cite{larson1998catalina}, Pan-STARRS \cite{kaiser2002pan}, and the Lincoln Near-Earth Asteroid Research project (LINEAR) \cite{stokes2000lincoln}, as well as space-based instruments such as the Wide-field Infrared Survey Explorer (WISE) mission \cite{wright2010wide}. Many of these asteroids are relatively small, but some are kilometre-sized and a few are predicted to closely encounter Earth or other planets in the near future. More encounter candidates are likely to be discovered by new efforts such as the Large-aperture Synoptic Survey Telescope (LSST) or the Las Cumbres Observatory (LCO). Light curve analysis has been performed on many asteroids to study their surface properties, but observations that are sensitive to the interior asteroid structure are rarer.

However, tidal torque acting on an asteroid and perturbing its angular velocity during an encounter presents an opportunity to constrain the interior mass distribution. Such angular velocity perturbations have been observed and used to extract asteroid properties in several cases, including for the 2013 encounter of (367943) Duende with Earth \cite{MOSKOVITZ2020113519, benson2020spin}, and asteroid binaries (3905) Doppler and (617) Patroclus respectively. Altered rotational states altered by Solar tidal torque have been excluded for 1I/2017 U1 `Oumuamua \cite{KWIECINSKI2018170}.

The upcoming 2029 encounter of 99942 Apophis is of particular interest and much theoretical work has been devoted to accurate estimates of its orbital and physical properties and predictions of how its angular velocity will be affected by tidal torque \cite{yu2014numerical, hirabayashi2021finite,valvano2022apophis, Lee2022Apophis, souchay2014rotational, souchay2018changes}. This research builds on a larger body of theoretical work. The gravitational two-body system has been studied in the context of tidal torque to different orders and with several different methods \cite{paul88, SCHEERES2000106, ashenberg07, BOUE2009750, HouMar2017}. Further studies showed that the tidal torque, observed through angular velocity perturbations, is sensitive to asteroid interior density distribution \cite{Naidu_2015, Makarov2022ChaosOO, RICHARDSON199847, scheeres2004evolution}. However, density distribution features beyond the moment of inertia ratios have not yet been extracted for any asteroid encounters. More research is needed to study in what cases these effects are observable, and what factors generally inhibit observation of these new features.

In this paper, we address this question by deriving a new equation of motion for an asteroid encounter system which is quick to compute and valid to arbitrary order. In section \ref{sec:methods}, we describe a simulation used to integrate the equations of motion and produce synthetic data of angular velocity over time, followed by a Markov Chain Monte Carlo (MCMC) fit process which  extracts physical characteristics of the asteroid (specifically, density moments) from the fit data. We then describe three methods to generate full density distributions from the density moments. Section \ref{sec:results} demonstrates the efficacy of this spin-data-to-density-distribution pipeline and explores its key properties. Finally, in section \ref{sec:discussion}, we measure the uncertainties of the best-fitting parameter estimates as a function of various encounter properties to determine which properties most inhibit accurate extraction of density distributions. We also test the density distribution extraction methods on several sample asteroids.

\section{Methods}
\label{sec:methods}

Only some aspects of an asteroid's density distribution affect tidal torque interactions. These are the asteroid ``length'' 
\begin{equation}
  a_\mathcal{A}^2 = \frac{1}{\mu_\mathcal{A}} \int_\mathcal{A} d^3 r \rho_\mathcal{A}(\bm r) r^2
  \label{eqn:am}
\end{equation}
and (complex) ``density moments''
\begin{equation}
  K_{\ell m} = \frac{1}{\mu_\mathcal{A} a_\mathcal{A}^\ell} \int_\mathcal{A} d^3 r \rho_\mathcal{A}(\bm r) R_{\ell m}(\bm r),
  \label{eqn:klm}
\end{equation}
both of which are constant and only need to be computed once to generate data. Here, $\rho_\mathcal{A}(\bm r)$ is the asteroid density distribution, $\mu_\mathcal{A}$ is the asteroid mass, and $R_{\ell m}$ are the regular solid spherical harmonics (see appendix \ref{app:eom} for details). They produce a tidal torque of 
\begin{equation}
  \begin{split}
  \bm \tau = & G\frac{\mu_\mathcal{A}\mu_\mathcal{B}}{2}\left[\sum_{\ell m} a_\mathcal{B}^\ell J_{\ell m} \sum_{\ell' m'}a_\mathcal{A}^{\ell'}S^*_{\ell+\ell', m + m'} (\bm D) (-1)^{\ell'}\right.\\
  \times & \left.\sum_{m''=-\ell'}^{\ell'} \sqrt{\frac{(\ell'-m'')!(\ell'+m'')!}{(\ell'-m')!(\ell'+m')!}}  \mathcal{D}^{\ell'}_{m'm''}(\alpha, \beta, \gamma)^* \right. \\
  \times & \Big((i\unit x - \unit y)(\ell'-m''+1)K_{\ell',m''-1} \\
  & +(i\unit x+\unit y)(\ell'+m''+1)K_{\ell',m''+1}+2im''\unit z K_{\ell'm''}\Big) \Bigg],
  \end{split}
  \label{eqn:tidal-torque}
\end{equation}
where $\bm D$ is the position of the asteroid; $\alpha$, $\beta$, and $\gamma$ are $z-y-z$ Euler angles expressing the orientation of the asteroid; $\mu_\mathcal{B}$, $a_\mathcal{A}$, and $J_{\ell m}$ are the mass, length, and density moments of the central body respectively; and $S_{\ell m}$ are the irregular solid spherical harmonics. Equation \ref{eqn:tidal-torque} is derived in appendix \ref{app:eom}, assuming a rigid asteroid and no third-body perturbations. Throughout the paper, we refer to the ``inertial frame'' (the frame in which the orbit is fixed) and the ``body-fixed frame'' (the frame in which the asteroid is fixed), which are also defined in appendix \ref{app:eom}.

In section \ref{sec:sim}, we describe the simulation used to integrate the equations of motion. In section \ref{sec:uncertainty}, we define the uncertainty model used to add noise to the data, and we describe how posterior probability distributions (PPDs) for density moments were extracted from spin-pole data in section \ref{sec:fit}. Finally, we outline a general model to extract density distributions from density moments in section \ref{sec:general-density} and define specific model implementations in section \ref{sec:specific-density}.

\subsection{Simulation design}
\label{sec:sim}

We built a publicly accessible, custom simulation in C++ to produce angular velocity data as a function of time. This simulation requires as initial data (1) the orbital parameters of the asteroid $r_p$ (perigee distance) and $v_\infty$ (hyperbolic excess velocity); (2) the cadence of angular velocity observation $\Delta t$; (3) the central body moments $J_{\ell m}$, mass $\mu_\mathcal{B}$, and radius $a_\mathcal{B}$; (4) the initial asteroid angular velocity in the inertial frame $\bm \Omega_0$; (5) the asteroid length $a_\mathcal{A}$; and (6) the asteroid's density moments $K_{\ell m}$ and initial Euler angle $\gamma_0$. All parameters except (6) are assumed to be known to high accuracy. One can imagine that $a_\mathcal{A}$ is approximated by light-curve analysis, but if not, it is still necessary to fix $a_\mathcal{A}$ or else the values of $K_{\ell m}$ are degenerate with $a_\mathcal{A}$.

We further assume that the asteroid is initially not tumbling. Thus, the angular velocity is aligned with a principal axis (assumed to be $\unit z$, which maximizes moment of inertia). This sets $\beta = 0$ and we can further choose $\alpha = 0$. Thus, only the Euler angle $\gamma_0$ is necessary to provide initial data for the simulation.

We begin our simulation at $D = 10 r_p$, with velocity predicted by Kepler's laws given by the orbital parameters. Since the leading order of the equations of motion is $\ell' = 2, \ell = 0$, this corresponds roughly to a torque of $10^{-3}$ times the maximum torque at perigee. Unless otherwise indicated, the simulation is terminated at $D=10 r_p$ as well.

With the simulation inputs specified, the equations of motion are integrated via the Runge-Kutta fourth order method, with a variable time step
\begin{equation}
  dt = dt_\text{min} + 10^{-3}(dt_\text{max} - dt_\text{min}) \brackets{\parens{\frac{D}{r_p}}^3 - 1}.
\end{equation}
The parameters $dt_\text{max}$ and $dt_\text{min}$ (20 and 10 seconds respectively) were chosen such that the numerical integration error was $\sim$100 times the floating point error, and that neighbouring values of $K_{\ell m}$ yielded significantly different spin pole data compared to floating point error. The data used to choose this $dt$ was obtained using the reference asteroid configuration, described in appendix \ref{app:reference-configs}.

\subsection{Uncertainty model}
\label{sec:uncertainty}

To add noise to data generated via the above simulation, we use the following uncertainty model. Each asteroid spin vector $\bm \Omega$ is assumed to be uncorrelated with other spin vectors, and we model uncertainty in the orientation and in the period as also uncorrelated. Consider a true spin vector $\bm \Omega^*$. For the sake of description, we work in coordinates in which $\bm \Omega^* \parallel \unit z$. Then, expressing the observed spin vector $\bm \Omega$ in spherical coordinates, we draw the polar angle from a normal distribution with standard deviation $\sigma_\theta$ centred on zero and the azimuthal angle from a uniform distribution. We also draw the ratio $\rho=\Omega/\Omega^*$ from a log-normal distribution centred on one, with width $\sigma_\rho$. This width is related to the uncertainty in asteroid period $\sigma_P$ by $\sigma_P=P \sigma_\rho$ for small $\sigma_\rho$. Explicitly, the probability density function (PDF) of $\rho$ is 
\begin{equation}
  P(\rho) = \frac{1}{\rho\sqrt{2\pi \sigma_\rho^2}} \exp\parens{-\frac{\ln^2\rho}{2\sigma_\rho^2}}.
\end{equation}
See figure \ref{fig:uncertainty-model} for an illustration of the uncertainty model.

\begin{figure}
  \centering
  \begin{tikzpicture}
  \draw[-{Latex[length=3mm]}] (0, 0) -- (-4, 0) node[anchor=east] {$\unit X$};
  \draw[-{Latex[length=3mm]}] (0, 0) -- (1, -1.5) node[anchor=west] {$\unit Y$};
  \draw[-{Latex[length=3mm]}] (0, 0) -- (0, 4) node[anchor=south] {$\unit Z$};

  \draw[line width=0.5mm, -{Latex[length=3mm]}] (0, 0) -- (-3, 3) node[anchor=south] {$\bm \Omega^*$};
  \draw[dashed] (-3, -2) -- (-3, 3);
  \draw[dashed] (-3, -2) -- (0, 0);

  \draw[-{Latex[length=3mm]}] (0, 0) -- (-1.25, 3.35) node[anchor=south west] {$\bm \Omega$};
  \draw[rotate around={-50:(-3,3)}] (-3,3) ellipse (1.4 and 2);
  \draw (-3, 3) arc (110:93:6);
  \draw (-2, 3.5) node[anchor=center] {$\theta$};
  \draw (-0.5, 2.5) node[anchor=center] {$\rho \Omega^*$};

  \end{tikzpicture}
  \caption{Diagram in the inertial frame of the uncertainty model used to define the probability that the true spin vector $\bm \Omega^*$ should be observed as $\bm \Omega$. The parameter $\theta$ is drawn from a Gaussian with width $\sigma_\theta$, and $\rho$ is drawn from a log normal distribution with width $\sigma_\rho$.}
  \label{fig:uncertainty-model}
\end{figure}

The log likelihood resulting from this uncertainty model is (excluding additive constants)
\begin{equation}
  \begin{split}
  \ln \mathcal{L} = -\frac{1}{2}\sum_{i = 0}\Bigg[&\frac{\cos^{-1} (\bm \Omega_i^* \cdot \bm \Omega_i/(\Omega_i^* \Omega_i))^2}{\sigma_\theta^2}\\
  &+\frac{\ln \parens{\Omega_i /\Omega_i^*}^2}{\sigma_\rho^2} + 2\ln\frac{\Omega_i}{\Omega_i^*}\Bigg].
  \end{split}
  \label{eqn:log-likelihood}
\end{equation}
where $\Omega_i$ is the $i$th spin vector in the data set.

This model was chosen because it separates spin pole and period uncertainty. Therefore, if one is more precisely determined by measurement, $\sigma_\theta$ and $\sigma_\rho$ can be adjusted separately in accordance.

\subsection{Extracting density moments from spin data}
\label{sec:fit}
Given synthetic data, an Affine Invariant MCMC Ensemble sampler was used to generate PPDs from flat priors. We use the Python implementation \texttt{emcee} \cite{foreman2013emcee}. Our parameters were $\gamma_0$, $K_{20}$, $K_{22}$, and $K_{3m}$ (10 in total), and were bounded by $|\gamma_0| < \pi/4$, and bounds on $K_{2 m}$ given in equation \ref{eqn:parameter-bounds}. Note that $\gamma_0$ is degenerate with $\gamma_0 + \pi/2$ because both of these align the asteroid principal axes with the body-fixed axes. The other bounds were $|K_{3m}| < 1$. In general, spin data is most sensitive to $\gamma_0$ and $K_{2m}$, which we call the ``first-order parameters.'' We call $K_{3m}$ the ``second-order parameters.''

The MCMC was determined to converge when the fractional change in autocorrelation time (computed every 100 iterations) was one percent, and the number of iterations computed so far was more than 100 times the autocorrelation time. The MCMC fit also was set to terminate if more than $10^5$ iterations were run, but this only occurred for fits in which the data was too low-quality to resolve the parameters. $\sim 10^4$ iterations was often sufficient, which generally consumed about 7 hours of computation time on a super computer running 16 threads on 8 cores.

Before the MCMC was run, local minima in the likelihood were found via the Nelder-Mead algorithm implemented in \texttt{scipy} \cite{Gao2012}. Generally, only one local minimum existed, except when $K_{22}=0$ in which case rotational symmetry caused multiple values of $\gamma_0$ to be degenerate. Ensemble walkers were initialized near this local minimum, distributed by a Gaussian approximation of the likelihood, as determined via the Hessian of the likelihood at the minimum. Due to the high sensitivity of the angular velocity data to density moments, the minimization procedure sometimes failed to isolate the minimum likelihood. Therefore, a simpler simulation without the $K_{3m}$ terms of equation \ref{eqn:tidal-torque} was first used to minimize likelihood as a function of the first-order parameters $\gamma_0$ and $K_{2m}$, and then the full simulation was used to find the second-order parameters $K_{3m}$ with the first-order parameters fixed.

To further ensure convergence, we minimized with respect to data truncated after perigee at double the perigee distance. This cut-off was manually chosen. The minimum was then further refined by minimizing based on the full data, with the previous minimum as the initial estimate.

\subsection{Extracting density distributions from moments: general model design}
\label{sec:general-density}

The asteroid density distribution $\rho_\mathcal{A}(\bm r)$ is not uniquely constrained via tidal torque interactions because only the density moments $K_{\ell m}$ contribute to equation \ref{eqn:tidal-torque}. However, by making sufficient assumptions about the density distribution, we can nevertheless estimate $\rho(\bm r)$ from $K_{\ell m}$. We outline three different assumptions which constitute three models for extracting density distributions from density moments. All are (partially) linear to facilitate computational ease and uncertainty propagation. One of them can be chosen, or another created, depending on what features are desired to be represented in the density distribution.

All three of these models assume that $a_\mathcal{A}$ and the asteroid shape is known (for example, by light curve or radar data). Uncertainty on this shape estimate is assumed to be small. The models then produce a density distribution $\rho_\mathcal{A}(\bm r)$ with uncertainty from $K_{\ell m}$ propagated to uncertainty in $\rho_\mathcal{A}$. If the asteroid shape is not known, a fourth model defined in appendix \ref{app:find-surface} can be used to extract it assuming uniform density. In section \ref{sec:results}, we were forced to fix $a_\mathcal{A}$ rather than fit for it because $a_\mathcal{A}$ is degenerate with scaling $K_{3m}$. However, $a_\mathcal{A}$ contains information about the density distribution as shown by its definition in equation \ref{eqn:am}. It might therefore be unreasonable to assume that it is exactly determined by light-curve or radar analysis. If $a_\mathcal{A}$ is too uncertain to be used as a known quantity when extracting density distributions from the moments, then the following models are flexible enough that the $a_\mathcal{A}$ constraint can be removed.

Since the overall mass of the asteroid is not observable from tidal torque, we do not expect to measure $\rho$ in an absolute sense. Only the differences in $\rho$ across the body are measurable.

To set the shape of the asteroid, we assume that an indicator function $\mathds{1}(\bm r)$ has been determined such that $\mathds{1}(\bm r) = 1$ inside the asteroid and 0 outside, where $\mathds{1}(\bm r)$ is defined in some frame whose orientation with respect to Earth is known. Since the asteroid rotates around its centre of mass during observations, we assume that the location of the centre of mass is also known in this frame, so that we can set it to be the origin.

We define a new coordinate system, the ``hybrid frame,'' which coincides exactly with body-fixed frame at the initial orientation of the asteroid assuming that the fit result for $\gamma_0$ is perfectly accurate. The orientation of the hybrid frame with respect to the inertial frame is therefore exactly known, so that $\mathds{1}$ is also exact in the hybrid frame and the body-fixed frame aligns with the hybrid frame up to uncertainty in $\gamma_0$. We will solve for $\rho(\bm r)$ in the hybrid frame given fit results $K_{\ell m}$ (known in the body-fixed frame) and the fixed $a_\mathcal{A}$ (the same in both frames).

We fix $a_\mathcal{A}$ at the value assumed during the fit and set $\mu_\mathcal{A}$ equal to an arbitrary constant since the system is independent of the total asteroid mass. Thus, $K_{\ell m}$ and $a_\mathcal{A}^2$ are linear in $\rho(\bm r)$ by equations \ref{eqn:am} and \ref{eqn:klm}. Equation \ref{eqn:am} and the $K_{00}=1$ component of equation \ref{eqn:klm} can each be applied as constraints to enforce these choices of $a_\mathcal{A}^2$ and $\mu_\mathcal{A}$.

Suppose that we restrict the number of degrees of freedom of $\rho(\bm r)$ from infinity to $m$ by explicitly defining some function $\rho(\bm r, \bm \Theta)$ for an $m$-dimensional vector $\bm \Theta$ which contains the free parameters of $\rho$. We will leave the explicit definition of $\rho(\bm r, \bm \Theta)$ to the model descriptions below, but for now we assume that $\rho$ is linear in $\bm \Theta$; i.e.,
\begin{equation}
  \rho(\bm r, \bm \Theta) = \bm B(\bm r)^\dagger \bm \Theta
  \label{eqn:density-distro}
\end{equation}
for a $m$-dimensional vector $\bm B(\bm r)$ with adjoint $\bm B(\bm r)^\dagger$. ($\bm B$ need not be linear in $\bm r$.) Thus, $a_\mathcal{A}^2$ and $K_{\ell m}$ are linear in $\bm \Theta$. We further assume that the model describes a way to reverse the relationship between $K_{\ell m}$ or $a_\mathcal{A}^2$ and $\bm \Theta$ so that
\begin{equation}
  \bm \Theta = A \bm K
  \label{eqn:density-model}
\end{equation}
where $A$ is a matrix. Here, we have arranged $a_\mathcal{A}^2$ and $K_{\ell m}$ into a vector $K$ which we say has $n$ dimensions. The order of this arrangement is irrelevant as long as it is kept consistent.

To propagate uncertainties from $\bm K$ to $\bm \Theta$ to $\rho(\bm r)$, we need the covariance matrix $\Sigma_K$ for $\bm K$. Suppose that the hybrid frame is offset from the body-fixed frame by some small angle $\Delta \gamma$, which results from uncertainty in $\gamma_0$. Then
\begin{equation}
  K_{\ell m}^\mathrm{hybrid} = e^{-im\Delta \gamma}K_{\ell m}^\mathrm{body-fixed}.
  \label{eqn:body-fixed-to-hybrid}
\end{equation}
by equation \ref{eqn:ylm-rotation}. Since $K_{\ell m}^\mathrm{body-fixed}$ was obtained by an MCMC fit, a large set of PPD-distributed samples is available for $\Delta \gamma$ and $K_{\ell m}^\mathrm{body-fixed}$, and the covariance matrix $\Sigma_K$ can be computed statistically in the hybrid frame by applying equation \ref{eqn:body-fixed-to-hybrid} to the samples. Then, propagation of uncertainty guarantees that the covariance matrix of $\bm \Theta$ is $\Sigma_\Theta = A \Sigma_K A^\dagger$ and the density distribution and uncertainty on density distribution are equal to
\begin{equation}
  \rho(\bm r) = \bm B(\bm r)^\dagger A\bm K \qquad \sigma^2_\rho(\bm r) = \bm B(\bm r)^\dagger A \Sigma_K A^\dagger \bm B(\bm r).
  \label{eqn:unc-rho}
\end{equation}

The role of a model is therefore to restrict the space of valid density distributions by defining the $m\times n$-dimensional matrix $A$ and the $m$-dimensional vector $\bm B(\bm r)$ such that equations \ref{eqn:density-distro} and \ref{eqn:density-model} are true. Then the density distribution and its uncertainty are given by equation \ref{eqn:unc-rho}. Three examples of possible models are given below.

\subsection{Extracting density distributions from moments: specific models}
\label{sec:specific-density}

\subsubsection{The ``likelihood'' model}
\label{sec:likelihood}

A natural way to restrict the degrees of freedom of $\rho(\bm r)$ is by defining a likelihood function $\mathcal{L}$ on the density distribution and choosing the one distribution which maximizes $\mathcal{L}$ and exactly reproduces $\bm K$. We call this the ``likelihood'' model. This likelihood should not be confused with the likelihood of equation \ref{eqn:log-likelihood}, which was a function of the spin data, not the density distribution. The choice of $\mathcal{L}$ is arbitrary, but a Gaussian likelihood will produce a linear model.

To employ this likelihood method, we divide the asteroid into a square grid of $m \gg n$ elements, each of which is assumed to have uniform density $\rho_0+\Theta_i$, where $\rho_0$ is constant across the asteroid and $\Theta_i$ is a local deviation. This defines the model function $\bm B(\bm r)$, which is zeroed in all components except for the $i$th, where $i$ is the index of the grid element that contains $\bm r$.

We use a likelihood of 
\begin{equation}
  \mathcal{L}(\bm \Theta) = \prod_{i=1}^m \frac{1}{\sqrt{2\pi \sigma^2}} \exp\parens{-\frac{\Theta_i^2}{2 \sigma^2}}
\end{equation}
with free parameters $\mu$ and $\sigma$. These parameters do not affect the location of the maximum, so we do not define them. Given this likelihood, the log likelihood is proportional to $-|\bm \Theta|^2$. Minimizing the norm of $\bm \Theta$ is therefore equivalent to finding the maximum likelihood.

Putting aside the problem of minimizing the norm, we use the linearity of equations \ref{eqn:am} and \ref{eqn:klm} to write $\bm K = M \bm \Theta$, where the $i$th entry of every row of the $n\times m$ matrix $M$ is the integral presented in equation \ref{eqn:am} or \ref{eqn:klm}, evaluated over the $i$th finite element. We want to solve $\bm K = M \bm \Theta$ for $\bm \Theta$ to match equation \ref{eqn:density-model}, which we do via the Moore-Penrose inverse. Since $m>n$, the Moore-Penrose inverse of $M$ is
\begin{equation}
  A=M^+ = M^\dagger(MM^\dagger)^{-1}.
  \label{eqn:mpi-underdetermined}
\end{equation}
The vector $\bm \Theta=M^+\bm K$ is guaranteed to solve $\bm K = M \bm \Theta$, and by the properties of the Moore-Penrose inverse, this $\bm \Theta$ also happens to minimize the norm of all possible $\bm \Theta$ that satisfy the equation. Thus, defining $A=M^+$ also minimizes $\mathcal{L}$ and fully defines the model.

This model is fast to compute; assuming fast matrix multiplication, the matrix inversion of equation \ref{eqn:mpi-underdetermined} is fast because $MM^\dagger$ is an $n$-dimensional square matrix, which is very small compared to the number of finite elements $m$.

\subsubsection{The ``harmonic'' model}
\label{sec:harmonic}

We now explore a model that seeks to restrict the space of allowed density distributions a different way: we allow only the density distributions with $\nabla^2 \rho = 0$ inside the asteroid (the harmonic distributions). We  call this  the ``harmonic'' model. Any harmonic function can be expanded as 
\begin{equation}
  \rho(\bm r) = \sum_{\ell m} \Theta_{\ell m}\frac{R_{\ell m}^*(\bm r)}{a_\mathcal{A}^\ell} 
  \label{eqn:harmonic-rho}
\end{equation}
where the terms which lead to $\rho \rightarrow \infty$ at the origin have been removed and $\Theta_{\ell m}$ are free (complex) parameters. Since $\rho$ is real, we have $\Theta_{\ell m}=(-1)^m \Theta_{\ell,-m}^*$. By setting a maximum on $\ell$, we restrict the number of degrees of freedom to $(\ell_\mathrm{max}+ 1)^2$. Choosing the same maximum $\ell$ as the maximum $\ell$ for the $K_{\ell m}$ moments, we have $m=n-1$ coefficients $\Theta_{\ell m}$ which can be stacked into an $m$-dimensional vector $\bm \Theta$.

Inserting equation \ref{eqn:harmonic-rho} into equations \ref{eqn:am} and \ref{eqn:klm},
\begin{equation}
  K_{\ell' m'} = \sum_{\ell m} \frac{1}{\mu_\mathcal{A} a_\mathcal{A}^{\ell'} a_\mathcal{A}^\ell} \Theta_{\ell m} \int_\mathcal{A} d^3 r R_{\ell m}^*(\bm r) R_{\ell' m'}(\bm r).
  \label{eqn:harmonic-mat}
\end{equation}
This is an over-determined matrix equation $\bm K = M \bm \Theta$, where $M$ is an $n \times m$ matrix. The Moore-Penrose inverse can therefore be used again to find an inverse $A=M^+$ which yields approximately correct $\bm \Theta$ (approximate in that the norm of the error vector between $\bm K$ and $M \bm \Theta$ is minimized). However, the form of the inverse changes due to the equation being overdetermined:
\begin{equation}
  A=M^+ = (M^\dagger M)^{-1} M^\dagger .
  \label{eqn:mpi-overdetermined}
\end{equation}

In the special case where the asteroid is a sphere of radius $R$, the matrix defined by equation \ref{eqn:harmonic-mat} is diagonal. %, with entries 
% \begin{equation}
%   M_{\ell m; \ell' m'} = \frac{4\pi R^3}{\mu_\mathcal{A}} \frac{R^{2\ell}}{a_\mathcal{A}^{2\ell}} \frac{\delta_{\ell \ell'} \delta_{m m'}}{(4\ell^2 + 8\ell + 3)(\ell - m)!(\ell+m)!}.
%\end{equation}
A non-spherical perturbation will introduce small, off-diagonal entries of $M$. This suggests another interpretation of the physical meaning of $K_{\ell m}$; they are directly proportional through this form of $M$ and the matrix equation $\bm K = M \bm \Theta$ to the coefficients of the spherical harmonic expansion of the asteroid density in the case of a spherical asteroid.

\subsubsection{The ``lumpy'' model}
\label{sec:lumpy}

The above two simple models produce smooth density distributions. In this section, we define a more complicated model which identifies discrete regions differing from the overall density. We call this model the ``lumpy'' model.

Suppose the asteroid contains $N$ ``lumps'' of uniform mass $\mu_i$ displaced by distance $\bm x_i$ from the asteroid centre of mass and superimposed on a constant-density overall asteroid ``medium'' with known shape given by $\mathds{1}(\bm r)$. For simplicity, we assume that all $N$ regions are ellipsoids with $d$ independent axis lengths. For example, $d=3$ corresponds to an asymmetric ellipsoid, $d=2$ corresponds to a symmetric ellipsoid, and $d=1$ corresponds to a sphere. The model therefore has $3N$ positional degrees of freedom, along with $N$ degrees of freedom for $\mu_i$, $Nd$ shape degrees of freedom, and 0, $2N$, or $3N$ rotational degrees of freedom for $d=1$, 2, or 3 respectively. The medium adds one additional degree of freedom for its mass, but its position is determined by knowledge of the surface $\mathds{1}(\bm r)$. The total number of degrees of freedom are displayed in table \ref{tab:lump-dof}. They should be compared to the number of known asteroid parameters $K_{\ell m}$ and $a_\mathcal{A}^2$, which is $(\ell_\text{max}+1)^2+1$: 17 when the second-order density moments are known and 10 when only $K_{2m}$ are known. We assume that the degrees of freedom of this lumpy model are fewer than the number of $\bm K$, so that the model is overdetermined.

%Recall that $\mathds{1}(\bm r)$ is known in the hybrid frame where the origin is the centre of mass of the asteroid. The displacement of the shape's centroid from the centre of mass, called $\bm x_0$, is therefore observable from light curve analysis and therefore constrains $\bm x_i$ so that the asteroid centre of mass lies on the origin.

\begin{table}
  \centering
  \begin{tabular}{cc|ccc}
    \hline \hline
        &  & & $N$ &  \\
        &  & 1  & 2  & 3  \\ \hline 
        & 1& \cellcolor{black}\color{white} 6 & \cellcolor{gray}\color{white} 11 & \cellcolor{gray}\color{white} 16\\
    $d$ & 2& \cellcolor{black}\color{white} 9 & \cellcolor{gray}\color{white} 17 & 25 \\
        & 3& \cellcolor{gray}\color{white}  11 &  21 &  31 \\
    \hline \hline
  \end{tabular}
  \caption{Total degrees of freedom $D$ as a function of $N$, the number of lumps modelled, and $d$, the number of independent axis lengths considered for each lump. The configurations with $D \leq$ the 10 known parameters not including $K_{3m}$ are coloured black, and with $D \leq$ the 17 parameters including $K_{3m}$ parameters are coloured gray.}
  \label{tab:lump-dof}
\end{table}

The net $K_{\ell m}$ components of such an ensemble are
\begin{equation}
  K_{\ell m} = K_{\ell m}'^0 + \sum_{i=1}^N K_{\ell m}'^i \qquad a_\mathcal{A}^2 = a_0'^2 + \sum_{i=1}^N a_i'^2
  \label{eqn:klm-stack}
\end{equation}
where $K_{\ell m}^i$ and $a_i^2$ obey the same definition as $K_{\ell m}$ in equation \ref{eqn:klm} and $a_\mathcal{A}$ in equation \ref{eqn:am} respectively integrated over the $i$th lump, but we set the normalizing factors $1/(\mu_\mathcal{A} a_\mathcal{A}^\ell)$ and $1/\mu_\mathcal{A}$ equal to the values for the entire asteroid, not their counterparts for each lump. The zero-indexed parameters indicate the moments of the medium. The prime in equation \ref{eqn:klm-stack} denotes that the moments are calculated in the hybrid frame, with its origin at the asteroid centre of mass.

We can relate the primed moments to the moments calculated relative to each lump's centre of mass via the translation rules for solid spherical harmonics:
\begin{equation}
  \begin{split}
  & K_{\ell m}'^i = \frac{1}{a_\mathcal{A}^{\ell - \ell'}}\sum_{\ell' m'} (-1)^{\ell - \ell'} R_{\ell - \ell', m - m'}(\bm x_i) K_{\ell' m'}^i\\
  & a_i'^2 = a_i^2 + x_i^2 \frac{\mu_i}{\mu_\mathcal{A}}
  \end{split}
  \label{eqn:translate-klm}
\end{equation}
from Ref.~\cite{Gelderen1998TheSO}. The dummy indices $\ell', m'$ should only be summed over values in which $\ell-\ell' \geq 0$ and $|m-m'| \leq \ell - \ell'$. Here, $\mu_i$ is the added mass of lump $i$, while $K_{1m}=0$ and $K_{2m}$ incorporate the lump's orientation and moment of inertia ratios. Its volume is constrained by $a_i^2$. These values map onto an ellipsoid shape via equations \ref{eqn:ellipsoid-axes}, so that if $K_{\ell m}^i$, $a_i^2$, $\mu_i$, and $x_i$ are known, then the density distribution of the asteroid is known.

Fixing $\bm x_i$, equation \ref{eqn:translate-klm} is linear in $K_{\ell m}^i$ and $a_i^2$. Therefore, if we define $\bm \Theta$ to contain $K_{2m}^i$ and $a_i^2$ for all $i$ in some order, then equations \ref{eqn:translate-klm} and \ref{eqn:klm-stack} define a matrix equation $\bm K = \bm C + M \bm \Theta$ which can be solved by setting $A$ equal to the Moore-Penrose inverse of $M$ (equation \ref{eqn:mpi-overdetermined}, since $\bm \Theta$ is overdetermined). The $\bm C$ term of this equation came from the expansion of $K_{\ell m}'^0$ and $a_0^2$, which is written in terms of the already-known displacement of the asteroid surface from its centre of mass and the surface shape.

We use a non-linear process to choose $\bm x_i$ and $\mu_i$ so that $K_{\ell m}^i$ and $a_i$ can then be extracted via the linear process described in section \ref{sec:general-density}. We must incorporate the centre-of-mass and total-mass constraints
\begin{equation}
  \mu_0 \bm x_0 + \sum_{i=1}^N \mu_i \bm x_i = 0, \qquad \mu_0 + \sum_{i=1}^N \mu_i = \mu_\mathcal{A}.
  \label{eqn:lump-constraints}
\end{equation}
There are also additional constraints, such as that $\bm x_i$ should lie inside the asteroid, which can be enforced manually. Combining both constraints in equation \ref{eqn:lump-constraints}, we can eliminate $\mu_0$ and vary only $\mu_i$ and $\bm x_i$ for $i \geq 1$.

The matrix equation defining $M$ is overdetermined, but we would like it to have a solution nevertheless. We therefore solve for $\bm x_i$ and $\mu_i$ by minimizing the difference between the known density moments and the output of the model. This is done by minimizing
\begin{equation}
  |(M(\bm x_i, \mu_i) M^+(\bm x_i, \mu_i) - \mathds{1}) (\bm K - \bm C)|^2.\
  \label{eqn:lump-minimize}
\end{equation}
where $\mathds{1}$ is the identity matrix. We cannot define $\bm B(\bm r)$ such that $\rho(\bm r)$ is linear in $\bm B$, but uncertainty on $K_{\ell m}^i$ can still be evaluated as $\Sigma_\Theta$ and these can be converted into uncertainties in the dimensions and orientations of the lumps. If the resulting density distribution is somehow excluded (it predicts negative density distributions or the lumps extend outside the asteroid, etc.), then another minimum of equation \ref{eqn:lump-minimize} can be used, or another combination of $N$ and $d$ listed in table \ref{tab:lump-dof}.

\section{Results}
\label{sec:results}

\subsection{Density moment extraction}
\label{sec:example-fit}
To demonstrate our fit process, we present example results of the MCMC fit described in section \ref{sec:fit}, applied to encounter data of the asymmetric reference asteroid. The fit results were consistent with the true density moments, and produced consistent data.

\begin{figure*}
  \centering
  \includegraphics[width=0.6\textwidth]{figs/example-residuals.pdf}
  \caption{Data, best-fitting results, and residuals for a fit to synthetic data simulated for an asymmetric reference asteroid. Uncertainty bands are also shown.}
  \label{fig:example-residuals}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=0.85\textwidth]{figs/example-corner.png}
  \caption{PPDs extracted from synthetic encounter data for the asymmetric reference asteroid. Samples from the MCMC fit are shown as individual points, and the contours enclose 1, 2, and 3$\sigma$ confidence regions. True values are shown as blue lines.}
  \label{fig:example-corner}
\end{figure*}

Figure \ref{fig:example-residuals} shows the spin data with noise added. The best-fitting model is overlaid in the top pane and residuals are shown the bottom pane. Specifically, we show the model whose parameters are the means of the PPDs. Uncertainties are plotted on the residuals corresponding to the square root of the diagonal entries of the covariance matrix (correlations not included). The fit results are clearly consistent with the data.

Figure \ref{fig:example-corner} shows a corner plot of the ten fitted PPDs, marginalized to functions of one (histograms) or two (contours) variables. The true parameters are shown as lines. Note that the true parameters usually lie within 1 or 2$\sigma$ of the $\Delta K_{\ell m} = 0$, where $\Delta K_{\ell m}$ is the difference between the posterior $K_{\ell m}$ and the true $K_{\ell m}$. The PPDs are generally Gaussian and sometimes show correlation between parameters, but no continuous degeneracy occurs. We performed 48 independent minimizations of the likelihood before the MCMC fit began, each with an initial point chosen randomly in the parameter space. All converged to the same minimum, demonstrating that the model lacks discrete degeneracy as well.

\subsection{Density distribution extraction}
\label{sec:asym-density}
To demonstrate our three density distribution extraction models, we show density distributions and uncertainties extracted for the asymmetric reference asteroid in figure \ref{fig:uniform-density-results}. These distributions were made given the density moment PPDs extracted above, with $a_\mathcal{A}$ fixed. The surface is fixed as a triaxial ellipsoid with semi-axis lengths such that the true values of $K_{\ell m}$ are consistent with uniform density (equation \ref{eqn:ellipsoid-axes}).

\begin{figure*}
  \centering
  \includegraphics[width=0.3\textwidth]{figs/asym-ell-likelihood.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/asym-ell-harmonic.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/asym-ell-lumpy.pdf}

  \includegraphics[width=0.3\textwidth]{figs/asym-ell-unc-likelihood.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/asym-ell-unc-harmonic.pdf}\hfill\hspace{0.3\textwidth}

  \caption{Density distributions (\textit{top}) and uncertainties (\textit{bottom}) extracted by the likelihood (\textit{left}), harmonic (\textit{middle}), and lumpy with $d=N=1$ (\textit{right}) models from data generated for a uniform-density, asymmetric reference asteroid. Density is normalized so that mean density is one.}
  \label{fig:uniform-density-results}
\end{figure*}

Also in each pane of figure \ref{fig:uniform-density-results} is $\Delta K_{\ell m}$, the squared magnitude of difference between the fitted density moments and the density moments of the final distribution. This difference is non-zero due to numerical error in the likelihood model case, which is guaranteed to exactly reproduce $K_{\ell m}$. The harmonic and lumpy models have no such guarantee, however.

The likelihood and harmonic models produce nearly identical density distributions. This is generally true for uniform density asteroids, which have harmonic density distributions. Consequentially, $\Delta K_{\ell m}$ is low for the harmonic model. This similarity lends confidence to both models, because if one were unexpectedly biased towards certain distributions, the other would have to be equally biased in the same way in this case to produce the same distribution, which appears unlikely given that the models share little in common.

The distributions are non-uniform for the harmonic and likelihood models (driven by error in the estimates of $K_{3m}$), but the non-uniformity is limited. Note that most of the asteroid has density close to the mean density (one) in both cases, with densities of 40\% or 160\% of the mean attained only in small regions. The lumpy model yields a much more uniform distribution because it is designed to produce uniform lumps. However, it does not produce a density distribution consistent with the moments extracted from spin data (i.e., $\Delta K_{\ell m}$ is large). This is because the lump found extends outside the asteroid, violating the assumptions of the model. This is a sign that the lump found does not really exist inside the asteroid.

The size of uncertainty is such that the density distribution is consistent with uniform for the harmonic and likelihood models (within 1.5$\sigma$). Uncertainty is generally larger towards the edges of the distribution, both for these shapes and for the shapes shown later.

\section{Discussion}
\label{sec:discussion}

\subsection{Spin evolution}
\label{sec:spin-evolution}
In figure \ref{fig:example-data}, we present example spin data generated via our simulation. A population of one thousand asteroids with identical initial conditions except for $\gamma_0$, $K_{20}$, and $K_{22}$ were simulated on a close Earth encounter. The exact parameters used were the symmetric and asymmetric cases described in appendix \ref{app:reference-configs}. Bands containing 68.3\%, 95.5\%, and 99.7\% of the population's spin are shown, as is the spin of the reference asteroids in black.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/nominal-data-sym.pdf}
  \includegraphics[width=\linewidth]{figs/nominal-data-asym.pdf}
  \caption{Angular velocity data simulated for the symmetric (above) and asymmetric (below) reference asteroids. The true angular velocity evolution is shown as a black line. Also plotted is the deviation of the data for PPD-distributed perturbations to the asteroid shape (bands). Bands contain 68.3\%, 95.5\%, and 99.7\% of the 1000 simulations run.}
  \label{fig:example-data}
\end{figure}

The initial values of $\gamma_0$, $K_{20}$, and $K_{22}$ were distributed proportional to the PPDs extracted via the MCMC fit shown in section \ref{sec:example-fit}. The PPDs were widened by a factor of 1000 to make the band widths visible. Therefore, the scale of the bands in figure \ref{fig:example-data} have little meaning in an absolute sense, but they are meaningful when comparing two bands or two times in the same band.

The figure illustrates the sensitivity of spin data to asteroid density moments and $\gamma_0$; Before perigee, all asteroids had similar angular velocities, but after perigee the angular velocities of the population diverged. The spin data at perigee is informative and shows non-trivial behaviour. After perigee, the asteroid population tumbles according to torque-free precession with different oscillation periods. This dramatically increases the spread of angular velocities within the distribution and allows for more precise refinement of the density moments. The oscillation periods observable during the torque-free precession phase of the spin data (about 5 hours after perigee) can be used to constrain $K_{22}$ and $K_{20}$, while the phase is set by the asteroid dynamics near perigee and therefore constrains the higher-order terms.

\subsection{Dependence of parameter uncertainty on encounter properties}
\label{sec:fit-uncertainty}

In this section, we discuss how the precision of the best-fitting density moments is affected by various encounter properties. We call this precision ``posterior uncertainty,'' or $\sigma(K_{\ell m})$, defining 1$\sigma$  posterior uncertainty as the range containing 68.27\% of the PPD centred on the median. 2$\sigma$ posterior uncertainty is defined likewise for 95.45\% of the PPD. These are roughly equivalent to 1 or 2 times standard deviation of the PPD respectively. We use the configuration of the asymmetric reference asteroid (appendix \ref{app:reference-configs}) unless otherwise stated.

\afterpage{
  \begin{landscape}
    \begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figs/scan-all1.pdf}
    \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of (left to right) perigee, excess velocity, spin pole uncertainty, and period uncertainty. The vertical dashed line indicates the reference asteroid values. The red vertical lines indicate when $\sigma(K_{3m}) =0.01$.}
    \label{fig:scan-perigee}
    \label{fig:scan-vex}
    \label{fig:scan-theta}
    \label{fig:scan-rho}
    \end{figure}
  \end{landscape}
}

\afterpage{
  \begin{landscape}
    \begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figs/scan-all2.pdf}
    \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of (left to right) asteroid length, rotational period, observational cadence, and data gap at perigee. The vertical dashed line indicates the reference asteroid values. The red vertical lines indicate when $\sigma(K_{3m}) =0.01$.}
    \label{fig:scan-am}
    \label{fig:scan-cadence}
    \label{fig:scan-period}
    \label{fig:observation-gap}
    \end{figure}
  \end{landscape}
}

There is no clear threshold above which posterior uncertainty is ``too high.'' However, most asteroid distributions tested in this paper have $K_{3 m} \lesssim 0.01$. We therefore call $\sigma(K_{3 m}) \approx 0.01$ the ``posterior uncertainty threshold,'' beyond which $K_{3 m}$ are unresolved. This should be understood as a rule of thumb, not a precise statement.

\subsubsection{Orbital elements}
\label{sec:scan-orbit}
A Keplerian orbit is completely described by five parameters, but three describe the orbit's orientation with respect to the central body. They are therefore redundant with the orientation of the inertial frame and we do not investigate them here. We parametrize the remaining two parameters by the perigee distance $r_p$ and excess velocity $v_\infty$. Fits of the type described in section \ref{sec:fit} were run for many values of $r_p$ and $v_\infty$ and the 1 and 2$\sigma$ posterior uncertainties are displayed in figure \ref{fig:scan-vex}.

Figure \ref{fig:scan-vex} demonstrates that posterior uncertainty follows a slight trend for uncertainty to increase with $v_\infty$. This is likely due to the fact that larger $v_\infty$ leads to a faster and flatter orbit with less time spent close to the planet, where tidal torque is strongest. There are also comparatively large oscillations in the uncertainty, due to the orientation of the asteroid at perigee varying. (The asteroid is always simulated to start at the same orientation, but increasing $v_\infty$ decreases the time to perigee so that the asteroid enters this region of high torque at different orientations depending on $v_\infty$.) This explains why the oscillations have the same period for all parameters. Note that these oscillations are sometimes large enough to raise $\sigma(K_{3m}) > 0.01$, shown by the red vertical line.

The figure shows much stronger dependence of parameter uncertainty on perigee distance, as expected by the factor of $(a_\mathcal{A}/D)^{\ell'}$ present in equation \ref{eqn:tidal-torque} and mentioned in section \ref{sec:tidal-torque}. For $r_p \approx 6$ or 7 Earth radii, $\sigma(K_{30})$ reaches the 0.01 threshold. At slightly higher $r_p$, $K_{31}$ and $K_{32}$ become similarly unresolved, and at $r_p \approx 20$ Earth radii, $K_{33}$, the last second-order component, becomes unresolved. Even before this, at $r_p \approx 10$ Earth radii, the most uncertain parameter $K_{30}$ fills the prior distribution with uncertainty ranging from -1 to 1, visible by the sudden cut-off in uncertainty increase and the discontinuity of the $\sigma(K_{\ell m})$ curve there. The exact location of these limits will change if different encounter properties are used.

%Fitted to each of the curves in figure \ref{fig:scan-perigee} are power law uncertainties $\sigma \propto r_p^\alpha$. These fits were performed via the method of least squares, and all data with $\sigma > 0.7$ was removed due to its sensitivity to the arbitrarily-chosen boundary of the prior. The values of $\alpha$ are shown in table \ref{tab:scan-perigee-alpha}. These slope values express how much each parameter is dependent on $r_p$. It is observed that $\gamma_0$ is least dependent on $r_p$, with $\sigma(\gamma_0) \sim r_p^2$. The other two first-order parameters are much more strongly dependent, with $\sigma \sim r_p^{5.5}$. The second-order parameters have milder slopes between 3 and 4 (except for $K_{30}$) which is fortunate from the perspective of making precise observations, since lower $\alpha$ renders larger values of $r_p$ accessible to measuring $K_{3m}$.

% \begin{table}
%   \centering
%   \begin{tabular}{l|c}
%     \hline
%     Parameter & $\alpha$ \\ \hline
%     $\gamma_0$ & $2.05$ \\
%     $K_{22}$ & $5.47$ \\
%     $K_{20}$ & $5.47$ \\ \hline
%     $\Re K_{33}$ & $3.35$ \\
%     $\Im K_{33}$ & $3.37$ \\
%     $\Re K_{32}$ & $3.05$ \\
%     $\Im K_{32}$ & $3.27$ \\
%     $\Re K_{31}$ & $3.53$ \\
%     $\Im K_{31}$ & $4.02$ \\
%     $K_{30}$ & $5.75$ \\ \hline
%   \end{tabular}
%   \caption{Power law slope values for the dependence of parameter uncertainty on perigee distance $r_p$. Slope is defined by $\sigma \propto r_p^\alpha$.}
%   \label{tab:scan-perigee-alpha}
% \end{table}

The axes of figure \ref{fig:scan-perigee} show that parameters with large $m$ are more precisely determined than parameters with small $m$, as can be seen by comparing $K_{22}$ to $K_{20}$ and comparing $K_{33}$ to other $K_{3m}$ values. Large $m$ moments correspond to moments that control higher frequency fluctuations in density at the asteroid equator. This pattern of $\sigma(K_{\ell m})$ smaller for large $m$ is a general trend and will be seen in the following sections as well.

The very strong dependence of $\sigma(K_{\ell m})$ on $r_p$ makes this analysis most useful to extract second-order moments on close encounters (5-20 Earth radii). Fortunately, in the case of Earth, these encounters are also likely to have the best associated observational uncertainty when above the horizon due to their proximity. The first-order moments can still be extracted at much larger perigee distances in our model.

\subsubsection{Observational uncertainty}
\label{sec:scan-uncertainty}
Two parameters, $\sigma_\theta$ and $\sigma_\rho$, govern the observational uncertainty of the data set. These parameters are defined in section \ref{sec:uncertainty}; $\sigma_\theta$ represents the standard deviation of the angle between the true spin pole and the observed spin pole, while $\sigma_\rho$ represents the standard deviation of the ratio between the observed and true rotational velocities. Rather than explore the full space spanned by these two values, we fix one and allow the other to vary to better assess whether uncertainty in spin pole or uncertainty in period more strongly affects posterior uncertainty $\sigma(K_{\ell m})$. This dependence is displayed in figure \ref{fig:scan-theta}.

For the case in which $\sigma_\rho$ is held fixed and $\sigma_\theta$ is varied, the figure shows that the dependence of $\sigma(K_{\ell m})$ on $\sigma_\theta$ is linear. The $K_{30}$ parameter reaches the $\sigma(K_{30})\approx 0.01$ threshold at $\sigma_\theta = 0.032\ \text{radians} \approx 2^\circ$, whereas the other parameters reach the threshold between $8^\circ$ and $45^\circ$. The $K_{2m}$ uncertainties remain low, even for $\sigma_\theta \approx 1$, which is large enough that the spin pole has non-negligible probability to be observed in any direction.

If $\sigma_\rho$ is allowed to vary, then $K_{30}$ reaches the threshold at $\sigma_\rho \approx \num{2.8e-7}$, and the other parameters reach the threshold at $\sigma_\rho \approx 1.2-\num{8.6e-6}$. These correspond to uncertainty in period of roughly milliseconds to seconds. At large period uncertainty, the PPDs begin to fill the prior and the posterior uncertainty $\sigma(K_{\ell m}$) is no longer linear with $\sigma_\rho$ (this is especially visible in the $K_{30}$) case. Otherwise, posterior uncertainty is proportional to $\sigma_\rho$.

These data show that posterior uncertainty is extremely sensitive to observational uncertainty in period in our analysis. To precisely measure density moments, very accurate rotational period estimates would have to be made every cadence. However this analysis does not study the effect of collecting more data after the encounter, or the correlations between different angular velocity measurements. If these correlations were taken into account, or a more accurate uncertainty model used which took into account the asteroid's location in the sky or its distance to Earth, then the precision limit we find here will be affected.

\subsubsection{Asteroid shape}
\label{sec:scan-shape}

THe true values of $K_{\ell m}$ and $a_\mathcal{A}$, which quantify the asteroid's physical properties, affect the posterior uncertainties. Here, we only investigate the sensitivity of $\sigma(K_{\ell m})$ to the first-order parameters and $a_\mathcal{A}$. These $K_{2m}$ moments can also be viewed as the axes of a uniform density triaxial ellipsoid (equation \ref{eqn:ellipsoid-axes}).

In figure \ref{fig:scan-space-sigma}, we show the 1$\sigma$ posterior uncertainties as a function of $K_{20}$ and $K_{22}$, or alternatively $a/c$ and $b/c$. We use axis ratios rather than the values of $a$, $b$, and $c$ to remove the $a_\mathcal{A}$ dependence of equation \ref{eqn:ellipsoid-axes}. The figure shows large uncertainty in $\gamma_0$ for $K_{22}=0$, or $a/c=b/c$, because $K_{20}$ is rotationally symmetric around $\unit z$, and $\gamma_0$ is the initial orientation with respect to the $\unit z$ axis. The system then has no dependence on $\gamma_0$ when $K_{22}=0$. This induces degeneracy in the model which inflates uncertainties, not only in $\gamma_0$ but also the other components.

\begin{figure*}
  \includegraphics[width=0.3\textwidth]{figs/probe-space-theta-1-sigma.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/probe-space-theta-2-sigma.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/probe-space-theta-3-sigma.pdf}

  \includegraphics[width=0.3\textwidth]{figs/probe-space-ab-1-sigma.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/probe-space-ab-a-sigma.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/probe-space-ab-b-sigma.pdf}

  \caption{1$\sigma$ posterior uncertainty for first-order parameters $\gamma_0$, $K_{22}$, and $K_{20}$ (\textit{top row}) and $\gamma_0$, $a/c$, and $b/c$ (\textit{bottom row}). Also shown as black points are the reference asteroid shapes: symmetric (red circle) and asymmetric (black triangle).}
  \label{fig:scan-space-sigma}
\end{figure*}

To remove the inflated uncertainty, one could assume a rotationally symmetric asteroid, remove $\gamma_0$ as a parameter, and run a fit. For a nearly rotationally symmetric asteroid however, a new parametrization is necessary which does not contain the ill-constrained $\gamma_0$ parameter. This task is beyond the scope of this paper, so we mostly consider asymmetric asteroids throughout.

Figure \ref{fig:scan-space-sigma} also shows low uncertainty for highly asymmetric asteroids, where $b/c$ and $a/c$ are very different (i.e., when $|K_{22}|$ is large). Additionally, $\sigma(K_{20})$ and $\sigma(K_{22})$ decrease for large $|K_{20}|$, which corresponds to large axis ratios in the ellipsoid case.

Figure \ref{fig:scan-space-corr} displays the correlation between the first-order parameters. They show that $\gamma_0$ and $K_{22}$ are often correlated for asymmetric asteroids, while $\gamma_0$ and $K_{20}$ are usually not. This is expected as $K_{22}$ is dependent on the orientation of the asteroid and $K_{20}$ is not. They also show that $K_{22}$ and $K_{20}$ are usually correlated, and that $a/c$ and $b/c$ are highly correlated. The latter is expected due to the $1/c$ dependence.

\begin{figure*}
  \includegraphics[width=0.3\textwidth]{figs/probe-space-corr12.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/probe-space-corr13.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/probe-space-corr23.pdf}

  \includegraphics[width=0.3\textwidth]{figs/probe-space-ab-1b.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/probe-space-ab-1a.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/probe-space-ab-ab.pdf}

  \caption{Correlations between PPDs for first-order parameters $\gamma_0$, $K_{22}$, and $K_{20}$ (\textit{top row}) and $\gamma_0$, $a/c$, and $b/c$ (\textit{bottom row}).  Also shown as black points are the reference asteroid shapes: symmetric (red circle) and asymmetric (black triangle).}
  \label{fig:scan-space-corr}
\end{figure*}

Overall, the variation in the uncertainties on $K_{20}$ and $K_{22}$ (the first-order density moments) is present but largely smooth across their allowed parameter space, as is their correlation. It therefore seems reasonable to use the asymmetric asteroid shape as a stand-in for an unknown's asteroid shape when simulating an encounter, as we do in this paper. The uncertainty then can be expected to differ across other shapes by a factor of about two or less, as long as the degenerate, symmetric asteroid regime is avoided.

On the other hand, the posterior uncertainty of $K_{3m}$ is much more strongly dependent on asteroid length $a_\mathcal{A}$. Figure \ref{fig:scan-am} displays posterior uncertainty $\sigma(K_{\ell m})$ as a function of $a_\mathcal{A}$. As was mention in section \ref{sec:tidal-torque}, the $K_{2m}$ parameters are insensitive to $a_\mathcal{A}$ since the $a_\mathcal{A}^2$ term in $\bm \tau$ (equation \ref{eqn:tidal-torque}) cancels the $a_\mathcal{A}^2$ in the moment of inertia (equation \ref{eqn:moi}). The $K_{3m}$ uncertainty is strongly dependent on $a_\mathcal{A}$ for the same reason that it is dependent on $r_p$: the $(a_\mathcal{A}/D)^{\ell'}$ dependence of equation \ref{eqn:tidal-torque}. At $a_\mathcal{A} \lesssim 700$ m, $K_{30}$ becomes unresolved. The other parameters become unresolved for $a_\mathcal{A} \lesssim 300$ m, with the cut-off for $K_{33}$ being less than 100 meters. This $a_\mathcal{A}$ cut-off is quite sharp; figure \ref{fig:scan-am} shows steep decrease in $\sigma(a_\mathcal{A})$ for low $a_\mathcal{A}$, with an approximate functional form of $\sigma(a_\mathcal{A}) \sim e^{(a_\mathcal{A}/C)^p}$ for a scaling constant $C$ and some power $p < 0.1$. This sharpness indicates that the cut-off value of $a_\mathcal{A}$ is unlikely to change significantly if other parameters of the flyby are altered in a way that slightly increases or decreases $\sigma(K_{\ell m})$. It therefore appears that extracting precise $K_{30}$ from asteroids with $a_\mathcal{A} \lesssim 700$ m is unlikely with this analysis, and for $a_\mathcal{A} \lesssim 50$ m, all $K_{3m}$ will likely be unresolved.

For uniform density asteroids, large $a_\mathcal{A}$ is equivalent to large asteroid radius. In non-uniform density asteroids, large $a_\mathcal{A}$ can also be achieved by distributing the mass of the asteroid near the surface, because the $r^2$ term in the integrand of the definition of $a_\mathcal{A}$ causes the density of regions distant from the asteroid centre of mass to dominate $a_\mathcal{A}$.

\subsubsection{Cadence}
\label{sec:scan-cadence}

The time between observations of asteroid angular velocity, or cadence, may vary depending on the observational schedule of the observing telescopes and the path of the asteroid through the sky.  We measure how the posterior uncertainty $\sigma(K_{\ell m})$ varies with cadence ranging from two minutes to one hour in figure \ref{fig:scan-cadence}.

Figure \ref{fig:scan-cadence} displays little dependence of uncertainty on cadence $\Delta t$ for $\Delta t \lesssim 40$ min. We also see flaring of uncertainty for very large cadence, largely driven by the paucity of data points. However, uncertainty dramatically increases for many parameters at about $\Delta t = 30-40$ min, a time scale which is likely characteristic of the asteroid system. We name this rough cadence limit $T_\text{cad}$.

We expect $T_\text{cad}$ to be a function of two dynamical time scales of the system: the rotational period of the asteroid $P_\omega$ and the time spent near perigee $T_p$. The latter can be estimated by
\begin{equation}
  T_p \sim \frac{r_p}{v_\infty}\brackets{2\frac{\mu_\mathcal{B}}{r_pv_\infty^2}+1}^{-\frac{1}{2}}
  \label{eqn:tp}
\end{equation}
which is the ratio of the perigee radius to velocity at perigee. The exact choice of the formula of $T_p$ not obvious, and alternatives to equation \ref{eqn:tp} are possible. For the simulated asteroid, $P_\omega = 9$ hr and $T_p = 42$ min by this definition.

%We further study the effect of $P_\omega$ and $T_p$ on $T_\text{cad}$ in appendix \ref{app:cadence-tests}, and find that both affect $T_\text{cad}$ by roughly the same amount.

Figure \ref{fig:scan-cadence} shows that as long as $\Delta t < T_\text{cad}$ is achieved, the influence of cadence on $\sigma$ is minimal, but shorter cadence leads to lower uncertainties.

\subsubsection{Perigee gap}
\label{sec:scan-gap}
In certain circumstances, spin data might not be able to be captured for a close encounter at perigee. The asteroid might dip below the horizon, or it might pass too close to the sun to be observed. Generally, angular velocity data can be collected when the asteroid is distant from the central body, where torque is low. There, the angular velocity evolution is dominated by torque-free precession dictated by the moment of inertia components. That zero-torque data can still be used to fix $K_{20}$ and $K_{22}$ as in \cite{MOSKOVITZ2020113519}. However, $K_{3m}$ does not affect the precession periods (though they do affect the phase). We are therefore curious as to how our posterior uncertainties change due to lack of perigee data.

To test this, we mask the perigee of the counter by removing a duration $T$ of data centred on the perigee, where $T$ ranges from 0 to 3 hours. To prevent lack of precision on $K_{\ell m}$ induced by lower amounts of data for high $T$, we always cut 3 hr$-T$ from the data set, half from the beginning and half from the end, so that each data set produced for all $T$ has the same size. We then fit the same asteroid model to the cut data for all $T$ and plot posterior uncertainties $\sigma(K_{\ell m})$ in figure \ref{fig:observation-gap}.

Since torque is greatest at perigee, we expect that region of the data to contain the most information about $K_{\ell m}$, and therefore uncertainty should increase monotonically with $T$, which is seen in figure \ref{fig:observation-gap}. We also see that the first-order parameters are not as sensitive to $T$ as the second-order parameters, because $K_{2m}$ are additionally constrained by torque-free precession after perigee.

Most parameters show dramatically increased uncertainty in the $T \sim 1-2$ hr range. On the other hand, none of the uncertainties increase noticeably for $T \lesssim 1$ hr. Thirty minutes of dropped data is equivalent to fifteen dropped points for the simulated cadence of $\Delta t = 2$ minutes, showing that many data points can dropped from the data set at perigee before the uncertainty starts to increase.

Qualitatively, \ref{fig:observation-gap} shows similar dependence of $\sigma(K_{\ell m})$ on $T$ as on cadence $\Delta t$; they also both have cut-offs where uncertainty markedly increases, and both the $T$ and $\Delta t$ cut-offs have qualitatively similar shapes although they occur at different values of $\Delta t$ and $T$. This suggests that the factors that govern uncertainty due to cadence also may govern sensitivity to lack of data at perigee in a similar way.

\subsubsection{Initial spin pole}
\label{sec:scan-spin}

The tidal torque experienced by the asteroid is affected by the initial direction of asteroid spin $\bm \Omega_0$ both because spin sets the initial asteroid orientation up to $\gamma_0$ and because of the spin-dependence of the rotational equations of motion (equation \ref{eqn:omega-eom}).

In figure \ref{fig:scan-spin}, we display 1$\sigma$ posterior uncertainties as a function of the direction of $\bm \Omega_0$ mapped onto the unit sphere in the inertial frame. Our samples for $\bm \Omega_0$ were laid out on a Fibonacci sphere to ensure they were roughly evenly spaced (marked in figure \ref{fig:scan-spin-avg}). To highlight common features across the parameters, we also display the average 1$\sigma$ sensitivity in figure \ref{fig:scan-spin-avg}. The average is weighted such that the uncertainty map for each parameter contributes an equal amount (the weight of each map is set to one-tenth of the map's mean). This average map is presented in two different projections to allow data at $\unit Z$ to be read.

\begin{figure*}
  \centering
  \includegraphics[width=0.8\textwidth]{figs/spin-pole.png}
  \caption{$1\sigma$ uncertainties for the first-order parameters (\textit{top}) and second-order (\textit{bottom}) as a function of the initial direction of spin in the inertial frame. All maps are made in the Mollweide projection. The orange star indicates the reference spin pole. The red contours enclose regions where $\sigma(K_{3m}) \geq 0.01$.}
  \label{fig:scan-spin}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=0.3\textwidth]{figs/spin-pole-avg-mollweide.png}
  \includegraphics[width=0.3\textwidth]{figs/spin-pole-avg-polar.png}
  \caption{The weighted average of the uncertainties shown in figure \ref{fig:scan-spin}, in Mollweide (\textit{left}) and polar (\textit{right}) projections in the inertial frame. See text for a description of how the average was computed. Black dots indicate the Fibonacci-sphere-distributed locations of sample spin poles, and the orange star indicates the reference spin pole. The polar projection is centred at the south pole, or $-\unit Z$.}
  \label{fig:scan-spin-avg}
\end{figure*}

Certain alignments of the body-fixed frame to the inertial frame lead to special conditions on torque, as discussed in section \ref{sec:tidal-torque}. For example, $\bm z \parallel \unit Z$ and $\bm z \parallel \unit Y$ at perigee lead to $\bm \tau \parallel \unit z$ to first-order, and $\bm \tau \parallel \unit X$ at perigee leads to $\bm \tau = 0$ to first-order. We relate this to the initial direction of $\bm \Omega_0$, via the approximation that $\bm \tau$ is small until perigee. Then $\bm \Omega_0 \parallel \unit Y$ and $\bm \Omega_0 \parallel \unit Z$ both lead to $\bm \tau \parallel \unit z$, and $\bm \Omega_0 \parallel \unit X$ leads to $\bm \tau = 0$.

Figure \ref{fig:scan-spin-avg} shows area of increased uncertainty for $\bm \Omega_0 \parallel \unit Z$ and $\bm \Omega_0 \parallel \unit Y$, but not the $\unit X$ case. This indicates that $\bm \tau \parallel \unit z$ causes increased uncertainty. Physically, $\bm \tau \parallel \unit z$ only changes an asteroid's rotational period and does not cause it to tumble, eliminating the ability to discern moment of inertia ratios from zero-torque precession after the encounter. If $\bm \tau = 0$ to first-order, then second-order $\bm \tau$ and non-perigee $\bm \tau$ will dominate, which may increase precision to these usually non-dominant parameters and therefore not have the same increasing effect on $\sigma(K_{\ell m})$.

However, uncertainty does not vary by much more than a factor of two outside the imprecise regions of $\bm \Omega_0 \parallel \unit Z$ and $\bm \Omega_0 \parallel \unit Y$, though these regions are wide for some parameters. Within the imprecise regions, uncertainty can grow up to four times or more the uncertainty at other $\bm \Omega_0$ values, and can exceed the $\sigma(K_{3 m}) \approx 0.01$ threshold. The trends for are roughly consistent across parameters (figure \ref{fig:scan-spin}), leading to clearly visible imprecise regions in the average $\sigma(K_{\ell m})$ (figure \ref{fig:scan-spin-avg}).

\subsubsection{Rotational period}
\label{sec:scan-period}

We also study the effect of the initial rotational period of the asteroid $P_\omega$ on posterior uncertainty $\sigma(K_{\ell m})$. In figure \ref{fig:scan-period}, we show $\sigma(K_{\ell m})$ as a function of $P_\omega$ for a range of periods typical of NEOs. Like figure \ref{fig:scan-vex}, depicting the dependence of $\sigma(K_{\ell m})$ on $v_\infty$, figure \ref{fig:scan-period} shows small-scale variation in uncertainty due to the fact that varying the initial period changes the value of $\gamma$ at perigee, which affects uncertainty to a factor of about two. But a large-scale trend is also visible in many parameters. $K_{20}$ and $K_{22}$ show very large uncertainty for $P_\omega \lesssim 4$ hr because these fast-rotators tumble very little after perigee. This increases uncertainty on the $K_{2m}$ parameters, which are constrained by tumbling.

We expect that quickly rotating asteroids would not tumble because, for small $P_\omega$, the dynamical variables $\bm D$, $\bm \omega$, $\alpha$, and $\beta$ vary much smaller than $\gamma$. Approximating each variable as constant over one full rotation of $\gamma$, we can integrate the first-order contribution of $\bm \tau$ over $\gamma \in (0, 2\pi)$, which gives no secular, first-order torque to force the asteroid to tumble. However, this effect does not apply to the second-order parameters, since the integral over the second-order term of $\bm \tau$ does not vanish, as seen in the figure.

Another feature of figure \ref{fig:scan-period} is that $K_{\ell 0}$ is more uncertain at low $P_\omega$ than the other parameters. This is most visible in the figure for $K_{30}$. The cause is likely that $K_{\ell 0}$ cannot contribute to $\tau_z$ as shown in equation \ref{eqn:tidal-torque}. We already discussed that asteroids with small $P_\omega$ do not tumble, and since $\tau_x$ and $\tau_y$ are what induces tumbling, the most observable component of torque is therefore $\tau_z$, which $K_{\ell 0}$ do not affect.

The most severe effect of period on $\sigma(K_{\ell m})$ is in the low-period regime ($P_\omega \lesssim 5$ hr), but in this case, the most strongly affected parameters are $K_{2m}$, which are generally known better than $K_{3m}$. The effect on the imprecise parameters $K_{3m}$ is small, except for $K_{30}$. It therefore seems as though small-period asteroids are still candidates for observation, although high-period asteroids yield smaller uncertainty.

\subsubsection{Central body oblateness}
\label{sec:scan-oblateness}

In all the above studies, we assumed a spherical planet ($J_{\ell m} = 0$ for $\ell \geq 1$). By assumption that $\mu_\mathcal{B} \gg \mu_\mathcal{A}$ (so that the asteroid orbit's focus is the centre of mass of the central body), we have $J_{1m} = 0$. The effect of central body non-sphericity, then, is limited to the $J_{2m}$ terms and damped by a factor of $(a_\mathcal{B} / D)^2$. We expect these parameters to have small effect on the asteroid.

Here, we define oblateness as $\epsilon = (I_z - I_x)/(\mu_\mathcal{B} R_\mathcal{B}^2)$, where $I_{x,y,z}$ are the central body moments of inertia along the principal axes, and $I_x = I_y$. $R_\mathcal{B}$ is the true radius of the body (not $a_\mathcal{B}$ from equation \ref{eqn:jlm}).

$J_{\ell m}$ is defined in equation \ref{eqn:jlm} with respect to the asteroid orbit, not the principal axes of the central body. However, for an equatorial orbit, the central body principal axes coincide with the asteroid orbit frame and we may express $\epsilon$ simply in terms of $J_{\ell m}$ as $\epsilon = -10J_{20}/3$ and $J_{22} = 0$. For simplicity, we use this equatorial orbit case. Since an oblate ellipsoid is mirror-symmetric around all three axes, table \ref{tab:klm-symmetries} indicates that $J_{3m}$ are all zero. The next order of tidal torque is therefore $J_{4m}$, damped by an additional $(a_\mathcal{B}/D)^2$ factor, and non-ellipsoid corrections to the central body shape. We do not consider these extra factors.

\begin{figure}
  \centering
  \includegraphics[width=0.88\columnwidth]{figs/oblateness.pdf}
  \vfill
  \includegraphics[width=0.88\columnwidth]{figs/oblateness-differ.pdf}
  \caption{\textit{Top}: 1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs as a function of oblateness $\epsilon$ . Linear best-fitting lines to $\sigma(K_{2m})$ (black, dotted) are plotted. \textit{Bottom}: The difference between PPD means extracted from a zero-oblateness model and the true parameters given data with true oblateness $\epsilon_\text{true} \neq 0$. Also shown in both figures are the oblatenesses of reference Solar System bodies.}
  \label{fig:scan-oblateness}
\end{figure}

Given this conversion between $\epsilon$ and $J_{20}$, we analyze posterior uncertainty $\sigma(K_{2 m})$ of the first-order parameters as a function of $\epsilon$ across a reasonable range of central body oblatenesses based on those of Solar System planets \cite{paterLissauer2015}. These uncertainties are shown in the top pane of figure \ref{fig:scan-oblateness}, together with linear best-fitting curves. The figure demonstrates almost no dependence of $\sigma(K_{\ell m})$ on oblateness $\epsilon$, although posterior uncertainty does measurably decrease for oblate central bodies. The best-fitting lines match the uncertainties well, and they have slope of $[\Delta \sigma(K_{\ell m}) / \sigma(K_{\ell m})_{\epsilon=0}] / \Delta \epsilon = -0.06$ for $\gamma_0$, $-0.2$ for $K_{22}$, and $-0.3$ for $K_{20}$. The second-order parameters $K_{3m}$ likely depend on oblateness similarly, but fitting these parameters is computationally more expensive and we do not study them.

Note that if an encounter is executed around one of the non-Earth objects noted in figure \ref{fig:scan-oblateness}, $a_\mathcal{B}$ and $\mu_\mathcal{B}$ will change in addition to $\epsilon$. These two parameters also affect the posterior uncertainty (section \ref{sec:jupiter-earth}), so the figure does not show that encounters with other bodies have the same precision as encounters with Earth; only that the difference in oblateness between the two bodies is of little concern.

Given the small effect of $\epsilon$ on $K_{\ell m}$, it might be tempting to neglect the planetary oblateness when fitting $K_{\ell m}$ to data. However, the bottom pane of figure \ref{fig:scan-oblateness} demonstrates that doing so is invalid. This figure displays $K_{\ell m}$ as extracted by a fit assuming $\epsilon = 0$, but run on data generated with non-zero $\epsilon$. The difference between the PPD means and true parameters are shown. Posterior uncertainties are also shown as bands. The figure shows that even for low (Earth-scale) oblateness, the fit results are inconsistent with the true $K_{\ell m}$ values, since $\Delta K_{\ell m} = 0$ is not contained in the 2$\sigma$ band. This effect is much worse for large oblateness, growing to a difference on the order of $\mathcal{O}(100)\sigma$ for Jupiter's oblateness. Therefore, accurately modelling central-body oblateness to high precision is essential for accurate estimation of fit parameters. For non-equatorial orbits, with $J_{22} \neq 0$, we also expect $J_{22}$ to affect the accuracy of the fit results to a similar degree, with the additional requirement of using the correct asteroid orbital plane.

$J_{20}$, the parameter studied in this section, has a slightly more general definition than oblateness. If the planet has a moon, the integral defining $J_{20}$ (equation \ref{eqn:jlm}) can be extended to include this extra mass, though this can only be done when the asteroid never passes inside the moon's orbit. As an order-of-magnitude estimate for this effect, two spherical objects with masses and radii of Earth and the Moon, separated by one Lunar distance, and both lying in the orbital plane has a combined oblateness of $\epsilon = 0.82$. Extrapolating posterior uncertainties via the slopes of the best fit lines given earlier yields a reduction in $\sigma(K_{2m})$ by about 25\%. Furthermore, $J_{22}$ is non-zero for this case, which likely decreases posterior uncertainty even more.

This analysis suggests that large moons such as ours can improve fit quality, but further study of this effect is beyond the scope of this paper. Without a moon to inflate the oblateness of the central body, planetary oblateness does not significantly improve posterior uncertainty. However, correct representation of oblateness is essential to accurately estimate $K_{\ell m}$.

\subsection{Comparison of Jupiter and Earth encounters}
\label{sec:jupiter-earth}

If sufficiently accurate spin pole data can be detected for non-Earth encounters, it may be possible to extract density moments for encounters with larger planets. In this section, we run our reference asteroid through a Jupiter encounter to analyze the differences in uncertainty.

The physical parameters of the asteroid body are kept the same as the Earth encounter case (listed in appendix \ref{app:reference-configs}), as are the observational uncertainty and cadence. The orbit is adjusted for the Jupiter case by setting a perijove distance of $r_p=5$ Jupiter radii (compared to perigee radius $r_p=5$ Earth radii for the Earth encounter). The excess velocity does not strongly affect $\sigma(K_{\ell m})$ as shown in section \ref{sec:scan-orbit}, so we keep it at the reference value. % We perform two fits, each with a different value of $v_\infty$: in one case we keep the Earth-encounter value of $v_\infty = 6$ km s$^{-1}$, and in the other case we increase the velocity such that the eccentricity (and therefore the angle between the orbits' hyperbolic asymptotes) are the same for both encounters. The latter requires $v_\infty = \sqrt{G\mu_\mathcal{B}/r_p (e-1)}=32$ km s$^{-1}$. The rationale for these excess velocity values is that, for $v_\infty = 6$ km s$^{-1}$, we keep a physically typical encounter excess velocity at the expense of changing the orbit shape, whereas for $v_\infty =32$ km s$^{-1}$ the orbit shape is unchanged, although the asteroid speed is dramatically increased.
%Section \ref{sec:scan-orbit} found that $v_\infty$ does not affect the posterior uncertainties, and indeed we find very little difference between the two $v_\infty$ cases. We therefore do not distinguish between the two cases further.
The ratio between the posterior uncertainties in the Jupiter and the Earth encounters are shown in table \ref{tab:jupiter-uncertainty}. In all cases, the Jupiter posteriors are more uncertain than Earth posteriors.

\begin{table}
  \centering
  \begin{tabular}{c|cc}
    \hline 
    $K_{\ell m}$ & $\sigma(K_{\ell m})_\text{Jupiter}/\sigma(K_{\ell m})_\text{Earth}$\\ \hline 
    $\gamma_0$ & 1.6 \\
    $K_{22}$ & 2.3 \\
    $K_{20}$ & 11 \\
    $\Re K_{33}$ & 18 \\
    $\Im K_{33}$ & 18 \\
    $\Re K_{32}$ & 18 \\
    $\Im K_{32}$ & 18 \\
    $\Re K_{31}$ & 25 \\
    $\Im K_{31}$ & 10 \\
    $K_{30}$ & 53 \\ \hline
  \end{tabular}
  \caption{Ratio of posterior uncertainty for all density moments $K_{\ell m}$ between an Earth encounter and a Jupiter encounter with identical properties except for an increased perigee.}
  \label{tab:jupiter-uncertainty}
\end{table}

These uncertainty ratios can be understood as follows. The leading order of tidal torque is proportional to $\mu_\mathcal{A} / D^3$. If $D/a_\mathcal{B}$ (the ratio of the encounter distance to the central body radius) is roughly constant (as in this case, where $r_p/a_\mathcal{B}=5$), then $\mu_\mathcal{A} / D^3 \propto \rho_\mathcal{B}$ where $\rho_\mathcal{B}$ is the density of the central body. Therefore, little advantage is to be gained by looking for encounters of a massive planet in this sense. Since Jupiter is less dense than Earth, we expect that uncertainty in the first-order parameters $\gamma_0$ and $K_{2m}$ would be slightly worse than in the case of Earth, which is seen in table \ref{tab:jupiter-uncertainty}.

The second-order terms are damped by an additional factor of $a_\mathcal{A}/D$, which decreases if a massive central body is used. Since Jupiter is about 10 times larger in radius than Earth, we expect that the $K_{3m}$ terms are about ten times more uncertain than the $K_{2m}$ components, which is the case.

The $K_{\ell 0}$ components differ in that the posterior uncertainty increase for a Jupiter encounter over an Earth encounter is about five times greater for $K_{\ell 0}$ than other moments of the same $\ell$. In fact, $K_{30}$ essentially fills the prior. In section \ref{sec:scan-spin}, we noted that $K_{\ell 0}$ is particularly uncertain when the asteroid does not tumble after the perigee. In this case, the Jupiter encounter resulted in less tumbling than the Earth encounter, so the larger increase in uncertainty in $K_{\ell 0}$ shown in table \ref{tab:jupiter-uncertainty} is expected.

There are additional effects of central body mass which are not captured in this analysis. For example, encounters with massive planets are more plentiful, so that observation for a fixed period of time will lead to a larger number of observed encounters conducive to low-uncertainty moment extraction (large $a_\mathcal{A}$, small $r_p$, etc.). The distribution of $r_p$ in this encounter sample will also change; the ratio of the orbit impact parameter (the distance between the orbit asymptotes and the central body) to perijove distance is 
\begin{equation}
  \frac{b}{r_p} = \sqrt{1+2\frac{G\mu_\mathcal{B}}{r_p v_\infty^2}}.
\end{equation}
It was mentioned above that the observable perigee distance $r_p$ increases with $a_\mathcal{B} \sim \mu_\mathcal{B}^{1/3}$, so that $\frac{b}{r_p}$ is larger for more massive planets and therefore the same impact parameter leads to lower perijove. Other effects, such as a change in the physical properties of the encountering asteroids or decreased observational uncertainty due to the distance between Jupiter and Earth-based telescopes, may also affect the fit uncertainties. Which of these contradicting effects dominates is not immediately clear, and depends on the asteroid population near Jupiter and the observation method.

\subsection{Comparisons between density distribution models}
\label{sec:density-compare}

To test the properties of the three density models defined in section \ref{sec:specific-density} to extract density distributions from density moments, we simulate several asteroids with different shapes and density distributions on a close Earth encounter with the reference asteroid orbital and observational parameters. In all cases, the density moments are extracted via our fit process. Below, we compare the resulting distributions to understand the performance of the density distribution models.

\subsubsection{Uniform density asteroids}

We simulate several asteroids with uniform density distributions and ask whether a uniform distribution is recovered. This was already done in section \ref{sec:asym-density} for the asymmetric reference asteroid. Here, we perform the same process for a ``dumbbell'' (two spheres of equal radius conjoined such that the surface of one intersects the centre of the other), the symmetric reference asteroid, and a tetrahedron. We use these shapes as examples of potential shape types (i.e., a contact binary for the dumbbell, a polyhedron for the tetrahedron) and to demonstrate the density models' efficacy on very different shapes. The extracted distributions for all three models are shown in figure \ref{fig:uniform-density}.

\begin{figure*}
  \centering
  \includegraphics[width=0.3\textwidth]{figs/db-likelihood.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/db-harmonic.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/db-lumpy.pdf}

  \includegraphics[width=0.3\textwidth]{figs/sym-ell-likelihood.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/sym-ell-harmonic.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/sym-ell-lumpy.pdf}

  \includegraphics[width=0.3\textwidth]{figs/tet-likelihood.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/tet-harmonic.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/tet-lumpy.pdf}
  \caption{Density distributions extracted by the likelihood (\textit{left}), harmonic (\textit{middle}), and lumpy with $d=N=1$ (\textit{right}) models from data generated for a uniform-density asteroid. The asteroid shapes are, from top to bottom, a dumbbell, the symmetric reference asteroid, and a tetrahedron. Density is normalized so that mean density is one. The same distance scale is used in all figures.}
  \label{fig:uniform-density}
\end{figure*}

From figure \ref{fig:uniform-density}, it can be seen that the likelihood and harmonic models produce mostly the same density distributions like in the case of section \ref{sec:asym-density}. Consequentially, $\Delta K_{\ell m}$ is low for the harmonic model. Also like section \ref{sec:asym-density}, the dumbbell asteroid produces largely uniform distributions except in a few small regions, and the lumpy model generates clearly incorrect results indicating that no lump is present.

The symmetric and tetrahedral asteroids are less-well represented by the density distribution model results. In both cases, the likelihood and harmonic models produce sometimes-negative distributions, and non-uniformity is especially large in the tetrahedral case. This is not the fault of the density models, but rather because both of these asteroids have $K_{22} = 0$ and (as discussed in section \ref{sec:scan-shape}), this leads to degeneracy in $\gamma_0$ and inflated uncertainty in the density moments. In the tetrahedral case, the problem is worsened because $K_{20}=0$ as well, so that torque is dominated by the small and ill-constrained $K_{3m}$ components. To avoid negative density distributions, the likelihood model could be adjusted to use, for example, a multivariate log-normal distribution. However, the model would no longer be linear.

Uncertainty in each distribution was calculated but not shown. The size of uncertainty is such that the density distribution is very consistent with uniform at most points for the (asymmetric) dumbbell model (within $\sim 1\sigma$), while the symmetric reference and tetrahedral asteroids were generally within $\sim 2 \sigma$ of uniform. Uncertainty is generally larger towards the edges of the distribution, both for these shapes and for the shapes shown in the following sections.

If desired, the parameters of the models can be reduced to only fit the $K_{2m}$ density moments; this will produce exactly uniform distributions, since it is equivalent to assuming $K_{3m}=0$. However, doing so ignores the observational insight we gain from our fits to $K_{3m}$, so it is only valid when $K_{3m}$ are too poorly constrained to be used.

\subsubsection{Spherical asteroids}
\label{sec:spherical-density}
In this section, we test the sensitivity of the density distributions to incorrect estimates for the shape of the asteroid by assuming a spherical shape for the asymmetric and symmetric reference asteroids, with density distributions shown in figure \ref{fig:sphere-density}.

\begin{figure*}
  \centering
  \includegraphics[width=0.33\textwidth]{figs/asym-sph-likelihood.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/asym-sph-harmonic.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/asym-sph-lumpy.pdf}

  \includegraphics[width=0.33\textwidth]{figs/sym-sph-likelihood.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/sym-sph-harmonic.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/sym-sph-lumpy.pdf}\

  \caption{Density distributions extracted by the likelihood (\textit{left}), harmonic (\textit{middle}), and lumpy with $d=N=1$ (\textit{right}) models from data generated for a uniform-density asteroid. The original asteroid shapes were the reference asymmetric (\textit{top}) and symmetric (\textit{bottom}) ellipsoids. In both cases, the density distributions have been extracted assuming a spherical shape. Density is normalized so that mean density is one. The same distance scale is used in all figures.}
  \label{fig:sphere-density}
\end{figure*}

Both the likelihood and harmonic models are accurate in that $\Delta K_{\ell m}$ is low, and they produce similar distributions as in the previous section. However, now even the asymmetric asteroid (with low uncertainty in $K_{\ell m}$) produces negative densities. This is a signal that the shape model is incorrect, and indeed the density distribution allows us to determine exactly how the shape model is wrong. In the asymmetric ellipsoid case, the density distribution is small (even negative) at large $|z|$ (top and bottom of the figure). The density is large at large $|y|$ (front right and back left in the figure). Comparison with the asymmetric reference asteroid depicted in figure \ref{fig:uniform-density-results} shows that the low-density region is outside the true asteroid shape, and the high-density region is inside. In other words, the shape can be corrected by extending it where density is large and retracting it where density is low.

A similar statement is true for the symmetric ellipsoid case in that density is low near the poles, but this time the regions with large density are evenly distributed around the equator, indicating the symmetry of the original asteroid.

The lumpy model produces lumps so large that they dominate the asteroid, predicting uniform distribution with $K_{\ell m} = 0$ for $\ell \geq 1$. The error in density moments $\Delta K_{\ell m}$ is therefore very large, equal to the sum of $|K_{\ell m}|^2$ for the original asteroids, making it clearly a poor model.

\subsubsection{Nonuniform asteroids}
\label{sec:non-uniform-density}

We also test several non-uniform density asteroids and compare the generated density distributions for all three models to see if the non-uniformities are recovered. We consider two asteroids, one with mass concentrated in the middle, and one with mass concentrated around the edges. (Specifically, the mass distribution follows a spherically symmetric exponential $\rho(r) \propto e^{\pm r^2/a_\mathcal{A}^2}$). To prevent the problems seen in previous sections when a symmetric asteroid is simulated, we set the shape of the asteroids to the asymmetric reference asteroid. The extracted density distributions are shown in figure \ref{fig:non-uniform-density}

\begin{figure*}
  \centering
  \includegraphics[width=0.3\textwidth]{figs/in-likelihood.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/in-harmonic.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/in-lumpy.pdf}

  \includegraphics[width=0.3\textwidth]{figs/out-likelihood.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/out-harmonic.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/out-lumpy.pdf}\

  \caption{Density distributions extracted by the likelihood (\textit{left}), harmonic (\textit{middle}), and lumpy with $d=N=1$ (\textit{right}) models from data generated for non-uniform-density asteroids with mass concentrated in the middle (\textit{top}) and the edges (\textit{bottom}). Density is normalized so that mean density is one. The same distance scale is used in all figures.}
  \label{fig:non-uniform-density}
\end{figure*}

For these non-uniform asteroids, the likelihood and harmonic model do not coincide, as can be seen by comparing the density distributions and that $\Delta K_{\ell m}$ is much larger for the harmonic distribution (indicating that the true density distribution is not harmonic). However, the two models give qualitatively similar results in that the centre-weighted density distribution contains more mass in the centre of the asteroid shape for both the harmonic and likelihood models while the edge-weighted asteroid contains mass at the edges. Though these general trends in mass placement arrived at by both the likelihood and harmonic models are accurate, the actual density distributions are inaccurate because they predict negative densities. The difference of the model output from the true distribution is generally less than $2\sigma$, but can rise to $3\sigma$ in some regions.

Figure \ref{fig:non-uniform-density} shows that the lumpy model places one large lump of greater density than the surrounding medium for the centre-weighted asteroid and less density for the edge-weighted asteroid, as expected. Greater error is incurred in $\Delta K_{\ell m}$ in this lumpy model than for the harmonic and likelihood models due to the lumpy model's crudeness, but the lump gives a good qualitative description of the density distribution of the asteroid which is useful for interpreting the other two models. This fact demonstrates that it can be illustrative to run the models with reduced degrees of freedom for the sake of simplicity, even if doing so yields less accurate distributions.

The lumpy model has one caveat that has not yet been mentioned, which applies for $N=d=1$ when the density moments are redundant with the asteroid shape; specifically, when (1) the centre of mass of the asteroid is the centroid of the asteroid shape and (2) the density moments of the asteroid shape are proportional to the fitted density moments. (1) guarantees that the lump will be placed at the origin and therefore will not contribute to $K_{\ell m}$ for $\ell \geq 1$, and (2) guarantees that the mass of the asteroid shape will be chosen such that the shape density moments exactly match the fitted density moments. The lump radius will then be set to zero because the fitted density moments are already matched. This zero-radius-lump result will occur even if the true density distribution has a large-radius lump at the origin. The lump model will therefore fail to recognize the lump in the true density distribution.

An example of this caveat applied is an elliptical lump inside an elliptical asteroid similar to the ones above. This is why, in the centre-weighted and edge-weighted density distributions, we specifically chose a spherically symmetric density distribution for our ellipsoidal asteroid shape rather than a distribution that conforms to the asteroid shape.

\subsubsection{Lumpy asteroids}

Finally, we test an asteroid with a discrete change in density distribution (for example, a rock of high density), of the kind that the lumpy model is designed to detect. We place a spherical mass of density six times that of the surrounding medium inside the asymmetric reference asteroid, slightly displaced along $\unit y$. Both the mass and the medium have uniform density. The distributions extracted by the three model for this asteroid are shown in figure \ref{fig:blob-density}.

\begin{figure*}
  \centering
  \includegraphics[width=0.3\textwidth]{figs/blob-likelihood.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/blob-harmonic.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/blob-lumpy.pdf}

  \caption{Density distributions extracted by the likelihood (\textit{left}), harmonic (\textit{middle}), and lumpy with $d=N=1$ (\textit{right}) models from data generated for an asteroid with an off-centre lump. Density is normalized so that mean density is one. The same distance scale is used in all figures.}
  \label{fig:blob-density}
\end{figure*}

We see from figure \ref{fig:blob-density} that the harmonic and likelihood model fail to recover the presence or location of the lump. There appears to be a slight increase in density at roughly the correct location for both models, but this is a small effect. Like figure \ref{fig:non-uniform-density}, we see that the harmonic and likelihood models do not agree exactly, and the failure of the harmonic model to exactly reproduce $K_{\ell m}$ suggests that the true distribution is not harmonic.

The failure of these models to find lumps is expected, since $K_{2 m}$ and $K_{3 m}$ are only sensitive to large-scale non-uniformities in the density and cannot exactly resolve the edges of the lump. However, the lumpy model computes the location and density of the mass very well. Some inaccuracies are induced by the uncertainty in $K_{3m}$, which changes the minimum of equation \ref{eqn:lump-minimize} such that the lump is displaced by 9 m along $\unit y$ from where it should be. This in turn causes a 70 m difference between the extracted lump's radius and the true radius. Nevertheless, it seems that the lumpy model is capable of recreating the location of lumps (except for the caveat mentioned above when the density moment data is redundant with the asteroid shape). But the ability of the lumpy model to place a lump does not guarantee the lump to exist, as was seen in the above sections.

\section{Conclusions}

We derived a novel, arbitrary-order equation for the tidal torque experienced by an asteroid during an encounter with a planet of arbitrary shape and mass distribution. The tidal torque equation (equation \ref{eqn:tidal-torque}) revealed that the angular velocity of the asteroid over time depends strongly on $K_{\ell m}$ and the initial orientation. We then built a fast simulation for an asteroid encounter, truncating the equation at second-order, and used the simulation to extract first- and second-order density moments from synthetic spin pole data via a Markov Chain Monte Carlo fit. We observe the following general properties of the moments:
\begin{enumerate}
  \item The second-order moments $K_{3m}$ are roughly a factor of $~\frac{a_\mathcal{A}}{r_p}$ more uncertain than the first-order moments, where $a_\mathcal{A}$ is proportional to the asteroid radius and $r_p$ is the perigee distance. For our reference asteroid, this fraction was about $10^{-5}$.
  \item The second-order moments that measure small-scale variations in density (i.e., $|m|$ is large) are generally less uncertain. It is often possible to measure $K_{3|m|}$ for large $m$ even when $K_{30}$ is not resolved.
  \item Extracted density moments are generally more precise when the asteroid leaves the encounter tumbling.
  \item The oblateness of the central body cannot be ignored without generating incorrect density moment estimates.
\end{enumerate}

We assessed the posterior uncertainty generated by the fit process by adjusting various properties of an asteroid on an Earth encounter. The main conclusions are listed below.

\begin{enumerate}
  \item Highly precise rotational period data is required to produce precise second-order moments.
  \item Large $a_\mathcal{A}$ is required for $K_{3m}$ resolution, though the size of $a_\mathcal{A}$ has no effect on $K_{2m}$. The threshold on $a_\mathcal{A}$ for the reference asteroid was $\sim$700 m for $K_{30}$ and $\sim$50 m for $K_{33}$.
  \item A close encounter is vital for accurate second-order moment extraction, though some of the second-order moments can still be required for more distant encounters.
  \item This model will yield inflated uncertainties if the asteroid is rotationally symmetric around its rotating spin axis. In this case, a different parametrization should be used. 
  \item If the asteroid rotational period is too small, or if the initial angular velocity points in an inconvenient direction, the density moments will be poorly constrained.
  \item The cadence of observation should generally be shorter than about 30 minutes for the reference asteroid case. Gaps of data can appear, even at perigee, but not for more than about an hour.
\end{enumerate}

If some of these conditions are not met, the first-order $K_{2m}$ moments may still be resolvable.

Finally, we presented three fast, linear methods for extracting density distributions (together with uncertainties) from the density moments and the shape of the asteroid The extracted density distributions are largely consistent with the true distributions in all cases we tested. Slight adjustments to $K_{3m}$ dramatically change the overall distribution, so that they are not very precise. Indeed, the models sometimes predict negative density (though more complicated models could be made not to). Errors in the shape estimate of the asteroid also distort the density distribution, but the resulting distributions provide hints as to exactly where the shape is incorrect. Large-scale properties of the asteroid's true distribution, such as whether the mass is located in the middle or on the edges, are often observable. Two of the models presented (the harmonic and lumpy models) are flexible in their number of parameters, so that if only these large-scale properties are necessary, a small set of parameters can be chosen to prevent over-fitting.`'

\section*{Acknowledgements}

We warmly thank Emmanuel Jehin, Maxime Devogele, and Marin Ferrais for meeting with the authors to discuss this work and pointing out possible future initiatives. JTD also thanks the MIT UROP office for funding his work. This paper made substantial use of MIT Supercloud's facilities.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Data Availability}

The asteroid simulation, fit process, and density moment extraction code are available on \href{https://github.com/jack-dinsmore/asteroid-tidal-torque}{GitHub}.

%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%

\bibliographystyle{mnras}
\bibliography{asteroids.bib}

\input{appendices}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bsp	% typesetting comment
\label{lastpage}
\end{document}